Changes in Patterns of Geometric Sign Groupings as a Model of Social Dynamics Across Aurignacian Phases

### 2025-05-06 endofquarterpaper.qmd
edit

Variation in diversity of sign types among sites may be representative of variation in social arrangement or social context. Despite the tendency to conceive of hunter-gatherers as monolithic, previous studies have established a potential for variation in hunter-gatherer social arrangements [@Finlayson_Warren_2017; @Riches_1995; @Price_Brown_1985; @Lane_2017]. @Bloch_1977 drew upon ethnographic studies of hunter-gatherers to theorize a correlation between the amounts of social structure, instituted hierarchy, and ritual communication within a society. @Conkey_1985 built upon that theorized correlation to argue that artwork and other forms of material culture can serve as a form of ritual communication, and that an increase in diversity within those material cultures may suggest more elaborate or complex social contexts. Changes in sign diversity from site to site, particularly the larger grouping structure of broader versus restricted signs, may be representative of two different social arrangement types, such as, for example, a more hierarchical structure versus a less hierarchical structure.

@@

One factor that may influence the geographic structure of sites is the possibility that some sites functioned as aggregation sites. The aggregation/dispersion pattern is first defined by @conkley_agg as a type of hunter-gather settlement pattern, in which groups both break off in dispersed fragments and congregate in large groups, typically for various subsistence or social reasons. Aggregation sites, associated with high degrees of connectivity, are the places where those group fragments or individuals congregate, and some archaeological indicators of aggregation sites include proximity to certain environmental resources, as well as a more diverse artifact assemblage [@conkley_agg]. Previous studies have also identified evidence of aggregation/dispersion patterns among the early modern humans of the Upper Paleolithic [@Maher_Conkey_2019; @bahn1982inter; @Bourdier_2013; @Svoboda_2013; @bahn1982inter; @white1992rethinking; @Soffer_Adovasio_Enloe_Audouze_Zubrow_2010; @montet-white1994]. It is therefore possible that sites with multiple connections between groups [@fig-neighbor-net], such as La Ferrassie in the the Proto-Aurignacian, Grotte de la Verpillière I in the Early Aurignacian, and Gargas in the Evolved Aurignacian acted as aggregation sites or served a similar purpose.

@@

signs <-
  input_df %>% 
  select(
         site_name,
         # c(line:star)
         # we are looking at the more visually complex signs
         cross, 
         v, 
         star, 
         zigzag, 
         zigzagrow, 
         grid, 
         hashtag, 
         hatching, 
         anthropomorph,
         zoomorph, 
         vulva) %>% 
  group_by(site_name) %>% 
  summarise(across(where(is.numeric), ~ sum(.))) %>% 
  column_to_rownames("site_name")

>>>>

Variation in diversity of sign types among sites may be representative of variation in social arrangement or social context. Despite the tendency to conceive of hunter-gatherers as monolithic, previous studies have established a potential for variation in hunter-gatherer social arrangements [@Finlayson_Warren_2017; @Riches_1995; @Price_Brown_1985; @Lane_2017]. @Bloch_1977 drew upon ethnographic studies of hunter-gatherers to theorize a correlation between the amounts of social structure, instituted hierarchy, and ritual communication within a society. @Conkey_1985 built upon that theorized correlation to argue that artwork and other forms of material culture can serve as a form of ritual communication, and that an increase in diversity within those material cultures may suggest more elaborate or complex social contexts. If the signs served functions of ritual communication, then changes in sign diversity from site to site, particularly the larger grouping structure of broader versus restricted signs, may be representative of two different social arrangement types, such as, for example, a more hierarchical structure versus a less hierarchical structure.

@@

One factor that may influence the geographic structure of sites is the possibility that some sites functioned as aggregation sites. The aggregation/dispersion pattern is first defined by @conkley_agg as a type of hunter-gather settlement pattern, in which groups both break off in dispersed fragments and congregate in large groups, typically for various subsistence or social reasons. Aggregation sites, associated with high degrees of connectivity, are the places where those group fragments or individuals congregate, and some archaeological indicators of aggregation sites include proximity to certain environmental resources, as well as a more diverse artifact assemblage [@conkley_agg]. Previous studies have also identified evidence of aggregation/dispersion patterns among the early modern humans of the Upper Paleolithic [@Maher_Conkey_2019; @bahn1982inter; @Bourdier_2013; @Svoboda_2013; @bahn1982inter; @white1992rethinking; @Soffer_Adovasio_Enloe_Audouze_Zubrow_2010; @montet-white1994]. It is therefore possible that sites with multiple connections between groups [@fig-neighbor-net], such as La Ferrassie in the Proto-Aurignacian, Grotte de la Verpillière I in the Early Aurignacian, and Gargas in the Evolved Aurignacian acted as aggregation sites or served a similar purpose.

@@

signs <-
  input_df %>% 
  select(site_name,
         c(line:star)) %>% 
  group_by(site_name) %>% 
  summarise(across(where(is.numeric), ~ sum(.))) %>% 
  column_to_rownames("site_name")
